Interval Probability for Fuzzy Quantum Theories

### 2016-05-30 prob.tex
prob.tex After meeting...

As motivated above, an event $E_1$ may have an interval of probability
$[l_1,r_1]$. Assume that another disjoint event $E_2$ has interval
probability $[l_2,r_2]$, what is the interval probability of the event
$E_1 \cup E_2$? The answer is somewhat subtle: although it is possible
to use the sum of the intervals $[l_1+l_2,r_1+r_2]$ as the combined
probability, one can do find a much tighter interval if information
\emph{against} the event (i.e., information about the complement
event) is also taken into consideration. Formally, for a general event
$E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability
$\left[1-r,1-l\right]$ which we abbreviate $1\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtracts intervals. Given
a collection of intervals $\mathscr{I}$, an
$\mathscr{I}$--interval-valued probability measure is a function
$\pmeas : \events \rightarrow \mathscr{I}$ such
that~\cite{JamisonLodwick2004}:\footnote{Notice that the belief function defined by Dempster~\cite{Dempster1967}
and Shafer~\cite{Shafer1976} is a special case of the interval-valued
probability measures.}
\begin{itemize}
\item $\pmeas(\emptyset)=[0,0]$, 
\item $\pmeas(\Omega)=[1,1]$, 
\item $\pmeas\left(\Omega\backslash E\right)=1\tensor*[_{\mathscr{I}}]{-}{}\pmeas\left(E\right)$,
  and 
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i}}\pmeas\left(E_{i}\right)$,
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly intervals. 
\end{itemize}

@@

@@

\yutsung{The definition of independence is different from different
sources:
\begin{enumerate}
\item In Swart's~\cite{Swart2013}, given a $*$-algebra~$\mathscr{A}$,
two commuting sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are \emph{logically
independent} if for all probability measure~$\mu_{1}$ on $\mathscr{A}_{1}$ and
$\mu_{2}$ on $\mathscr{A}_{2}$, there is an unique probability measure~$\mu$
on the smallest sub-$*$-algebra of $\mathscr{A}$ containing both
$\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ such that $\mu\left(A_{1}
A_{2}\right)=\mu_{1}\left(A_{1}\right)\mu_{2}\left(A_{2}\right)$
for all $A_{1}\in\mathscr{A}_{1}$ and $A_{2}\in\mathscr{A}_{2}$. If two
sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are logically
independent, the smallest sub-$*$-algebra containing both $\mathscr{A}_{1}$ and
$\mathscr{A}_{2}$ is denoted by $\mathscr{A}_{1}\otimes\mathscr{A}_{2}$, and
its probability measure~$\mu$ is denoted by $\mu_{1}\otimes\mu_{2}$\\
Then, Swart proved the Bell inequality on logically independence $*$-algebras
with a product state which gives a clear physical meaning. However, this
only defines the independence of algebras instead of events.
\item In Yuan's~\cite{Yuan2012}, given a $\mu$ on a $*$-algebra~$\mathscr{A}$,
two sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are \emph{independent}
if $\mu\left(A_{1}A_{2}\right)=\mu\left(A_{1}\right)\mu\left(A_{2}\right)$
for all $A_{1}\in\mathscr{A}_{1}$ and $A_{2}\in\mathscr{A}_{2}$.
The same problem as the previous one: this defines the
independent of algebras, but not events. Morever, I haven't found a clear
physical meaning of this definition, yet\ldots
\item In order to understand the definition of independence in~\cite{GuehneKleinmannCabelloEtAl2010},
we need to first understand their notation because they focused on the
hidden-variable (HV) models. Let $\lambda$ is the HV, $A_{i}$ is
the measurement of the observable~$A$ at the position $i$ in the
sequence. For example, $A_{1}B_{2}C_{3}$ denotes the sequence of
measuring $A$ first, then $B$, and finally $C$. Given a fixed $\lambda$, we
assume the outcome of an observable is deterministic. For example, the
outcome of $B_{2}$ from the preceding sequence is denoted by $v\left(B_{2}\middle|A_{1}B_{2}C_{3}\right)$.
Then, the outcome of $A$ is \emph{independent} of whether $B$ is measured
before or after $A$ is $v\left(A_{1}\right)=v\left(A_{2}\middle|B_{1}A_{2}\right)$.\\
If the outcomes of compatible observables are independent for any
HV~$\lambda$, the HV model is non-contextual. (Notice that they have their own
definition of ``compatible observables'' for their HV model\ldots)\\
Compare to the previous definitions, this definition is appealing
in a sense that projection operators are special cases of observables,
and this definition has a physical meaning related to contextuality.
However, before adopting their definition, we need to reformulate
the definition to the usual quantum model....
\item For the definition we used last time, given a quantum probability
measure~$\mu:\events\rightarrow[0,1]$, two commuting projections
$P_{1}$ and $P_{2}$ are \emph{independent} if
$\mu\left(P_{1}P_{2}\right)=\mu\left(P_{1}\right)\mu\left(P_{2}\right)$.
We have two examples:
\end{enumerate}
}

\yutsung{\begin{example}[Independence of product state] Consider
the 2-qubit Hilbert space with a probability measure~$\mu\left(P\right)=\melement{\ps\ps}{P}$
with the following projections:
\[
\begin{array}{rcl}
P_{1}=\proj{00}+\proj{01},\quad P_{2}=\proj{00}+\proj{10},\quad P_{3}=\proj{10}+\proj{11},\quad P_{4}=\proj{00}\end{array}
\]
and their probabilities: 
\[
\begin{array}{rcl}
\mu\left(\mathbb{0}\right)=0,\quad\mu\left(P_{1}\right)=1/2,\quad\mu\left(P_{2}\right)=1/2,\quad\mu\left(P_{3}\right)=1/2,\quad\mu\left(P_{4}\right)=1/4\textrm{ .}\end{array}
\]
On one hand, $P_{1}$ and $P_{3}$ are complement to each other so
we have
\begin{eqnarray*}
\mu\left(P_{1}\right)+\mu\left(P_{3}\right) & = & \frac{1}{2}+\frac{1}{2}=1\\
\mu\left(P_{1}\right)\mu\left(P_{3}\right) & = & \frac{1}{4}\ne0=\mu\left(\mathbb{0}\right)=\mu\left(P_{1}P_{3}\right)
\end{eqnarray*}
Therefore, $P_{1}$ and $P_{3}$ commute, but are not independent. On
the other hand, $P_{1}$ also commutes with $P_{2}$, and
\begin{eqnarray*}
\mu\left(P_{1}\right)\mu\left(P_{2}\right) & = & \frac{1}{4}=\mu\left(P_{4}\right)=\mu\left(P_{1}P_{2}\right)
\end{eqnarray*}
so that $P_{1}$ and $P_{2}$ are independent.\qed\end{example}

\begin{example}[Independence of any states] In general, for any orthonormal
basis~$\left\{ \ket{\psi_{j}}\right\} _{j=0}^{3}$, we can always
pick 
\[
\begin{array}{rcl}
P_{1}=\proj{\psi_{0}}+\proj{\psi_{1}},\quad P_{2}=\proj{\psi_{0}}+\proj{\psi_{2}},\quad P_{3}=\proj{\psi_{0}}\textrm{ .}\end{array}
\]
With the probability measure~$\mu\left(P\right)=\melement{\psi}{P}$,
where $\ket{\psi}=\frac{1}{2}\sum_{j=0}^{3}\ket{\psi_{j}}$, their
probabilities are
\[
\begin{array}{rcl}
\mu\left(P_{1}\right)=1/2,\quad\mu\left(P_{2}\right)=1/2,\quad\mu\left(P_{3}\right)=1/4\textrm{ ,}\end{array}
\]
and we have
\begin{eqnarray*}
\mu\left(P_{1}\right)\mu\left(P_{2}\right) & = & \frac{1}{4}=\mu\left(P_{3}\right)=\mu\left(P_{1}P_{2}\right)\textrm{ .}
\end{eqnarray*}
In particular, the following four Bell states form an orthonormal
basis~\cite{CabelloPRL.86.2001}: 
\[
\ket{\phi^{\pm}} = \frac{1}{\sqrt{2}}\left(\ket{00}\pm\ket{11}\right),\quad
\ket{\psi^{\pm}} = \frac{1}{\sqrt{2}}\left(\ket{01}\pm\ket{10}\right)\textrm{ .}
\]
Their sum is also a Bell state~
\[
\ket{\psi}=\frac{1}{2}\left(\ket{\phi^{+}}+\ket{\phi^{-}}+\ket{\psi^{+}}+\ket{\psi^{-}}\right)=\frac{1}{\sqrt{2}}\left(\ket{00}+\ket{01}\right)\textrm{ .}
\]
Thus, $\proj{\phi^{+}}+\proj{\psi^{+}}$ and $\proj{\phi^{+}}+\proj{\psi^{-}}$
are independent projections with respect to the probability
measure~$\mu\left(P\right)=\melement{\psi}{P}$.\qed\end{example}

The definition of independence is interesting, and we definitely need
to discuss it when we want to discuss Bell's theorem and the Kochen-Specker
theorem. However, the definition diverges, and none of the source I have checked 
support our definition, yet. We may need to check more papers to find a source
to support our definition or to find the majority definition (if there is one)...}

@@

>>>>

As motivated above, an event $E_1$ may have an interval of probability
$[l_1,r_1]$. Assume that another disjoint event $E_2$ has interval
probability $[l_2,r_2]$, what is the interval probability of the event
$E_1 \cup E_2$? The answer is somewhat subtle: although it is possible
to use the sum of the intervals $[l_1+l_2,r_1+r_2]$ as the combined
probability, one can do find a much tighter interval if information
\emph{against} the event (i.e., information about the complement
event) is also taken into consideration. Formally, for a general event
$E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability
$\left[1-r,1-l\right]$ which we abbreviate $1\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtracts intervals. Given
a collection of intervals $\mathscr{I}$, an
$\mathscr{I}$--interval-valued probability measure is a function
$\pmeas : \events \rightarrow \mathscr{I}$ such
that~\cite{JamisonLodwick2004}:\footnote{Notice that the belief function defined by Dempster~\cite{Dempster1967}
and Shafer~\cite{Shafer1976} is a special case of the interval-valued
probability measures. Depending on how quantum interval probability goes, this
section may be specialized to Dempster and Shafer only.}
\begin{itemize}
\item $\pmeas(\emptyset)=[0,0]$, 
\item $\pmeas(\Omega)=[1,1]$, 
\item $\pmeas\left(\Omega\backslash E\right)=1\tensor*[_{\mathscr{I}}]{-}{}\pmeas\left(E\right)$,
  and 
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\tensor*[_{\mathscr{I}}]{\sum}{_{i}}\pmeas\left(E_{i}\right)$,
where $\tensor*[_{\mathscr{I}}]{\sum}{_{i}}[l_{i},r_{i}]=\left[\tensor*[_{\mathbb{R}}]{\sum}{_{i}}l_{i},\tensor*[_{\mathbb{R}}]{\sum}{_{i}}r_{i}\right]$.
We may drop the preposing $\mathscr{I}$ when summands are clearly intervals. 
\end{itemize}

@@

\subsection{Quantum Events}

@@

% \yutsung{The definition of independence is different from different
% sources:
% \begin{enumerate}
% \item In Swart's~\cite{Swart2013}, given a $*$-algebra~$\mathscr{A}$,
% two commuting sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are \emph{logically
% independent} if for all probability measure~$\mu_{1}$ on $\mathscr{A}_{1}$ and
% $\mu_{2}$ on $\mathscr{A}_{2}$, there is an unique probability measure~$\mu$
% on the smallest sub-$*$-algebra of $\mathscr{A}$ containing both
% $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ such that $\mu\left(A_{1}
% A_{2}\right)=\mu_{1}\left(A_{1}\right)\mu_{2}\left(A_{2}\right)$
% for all $A_{1}\in\mathscr{A}_{1}$ and $A_{2}\in\mathscr{A}_{2}$. If two
% sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are logically
% independent, the smallest sub-$*$-algebra containing both $\mathscr{A}_{1}$ and
% $\mathscr{A}_{2}$ is denoted by $\mathscr{A}_{1}\otimes\mathscr{A}_{2}$, and
% its probability measure~$\mu$ is denoted by $\mu_{1}\otimes\mu_{2}$\\
% Then, Swart proved the Bell inequality on logically independence $*$-algebras
% with a product state which gives a clear physical meaning. However, this
% only defines the independence of algebras instead of events.
% \item In Yuan's~\cite{Yuan2012}, given a $\mu$ on a $*$-algebra~$\mathscr{A}$,
% two sub-$*$-algebra $\mathscr{A}_{1}$ and $\mathscr{A}_{2}$ are \emph{independent}
% if $\mu\left(A_{1}A_{2}\right)=\mu\left(A_{1}\right)\mu\left(A_{2}\right)$
% for all $A_{1}\in\mathscr{A}_{1}$ and $A_{2}\in\mathscr{A}_{2}$.
% The same problem as the previous one: this defines the
% independent of algebras, but not events. Morever, I haven't found a clear
% physical meaning of this definition, yet\ldots
% \item In order to understand the definition of independence in~\cite{GuehneKleinmannCabelloEtAl2010},
% we need to first understand their notation because they focused on the
% hidden-variable (HV) models. Let $\lambda$ is the HV, $A_{i}$ is
% the measurement of the observable~$A$ at the position $i$ in the
% sequence. For example, $A_{1}B_{2}C_{3}$ denotes the sequence of
% measuring $A$ first, then $B$, and finally $C$. Given a fixed $\lambda$, we
% assume the outcome of an observable is deterministic. For example, the
% outcome of $B_{2}$ from the preceding sequence is denoted by $v\left(B_{2}\middle|A_{1}B_{2}C_{3}\right)$.
% Then, the outcome of $A$ is \emph{independent} of whether $B$ is measured
% before or after $A$ is $v\left(A_{1}\right)=v\left(A_{2}\middle|B_{1}A_{2}\right)$.\\
% If the outcomes of compatible observables are independent for any
% HV~$\lambda$, the HV model is non-contextual. (Notice that they have their own
% definition of ``compatible observables'' for their HV model\ldots)\\
% Compare to the previous definitions, this definition is appealing
% in a sense that projection operators are special cases of observables,
% and this definition has a physical meaning related to contextuality.
% However, before adopting their definition, we need to reformulate
% the definition to the usual quantum model....
% \item For the definition we used last time, given a quantum probability
% measure~$\mu:\events\rightarrow[0,1]$, two commuting projections
% $P_{1}$ and $P_{2}$ are \emph{independent} if
% $\mu\left(P_{1}P_{2}\right)=\mu\left(P_{1}\right)\mu\left(P_{2}\right)$.
% We have two examples:
% \end{enumerate}
% }
% 
% \yutsung{\begin{example}[Independence of product state] Consider
% the 2-qubit Hilbert space with a probability measure~$\mu\left(P\right)=\melement{\ps\ps}{P}$
% with the following projections:
% P_{1}=\proj{00}+\proj{01},\quad P_{2}=\proj{00}+\proj{10},\quad P_{3}=\proj{10}+\proj{11},\quad P_{4}=\proj{00}\end{array}
% and their probabilities: 
% \mu\left(\mathbb{0}\right)=0,\quad\mu\left(P_{1}\right)=1/2,\quad\mu\left(P_{2}\right)=1/2,\quad\mu\left(P_{3}\right)=1/2,\quad\mu\left(P_{4}\right)=1/4\textrm{ .}\end{array}
% On one hand, $P_{1}$ and $P_{3}$ are complement to each other so
% we have
% \mu\left(P_{1}\right)+\mu\left(P_{3}\right) & = & \frac{1}{2}+\frac{1}{2}=1\\
% \mu\left(P_{1}\right)\mu\left(P_{3}\right) & = & \frac{1}{4}\ne0=\mu\left(\mathbb{0}\right)=\mu\left(P_{1}P_{3}\right)
% Therefore, $P_{1}$ and $P_{3}$ commute, but are not independent. On
% the other hand, $P_{1}$ also commutes with $P_{2}$, and
% \mu\left(P_{1}\right)\mu\left(P_{2}\right) & = & \frac{1}{4}=\mu\left(P_{4}\right)=\mu\left(P_{1}P_{2}\right)
% so that $P_{1}$ and $P_{2}$ are independent.\qed\end{example}
% 
% \begin{example}[Independence of any states] In general, for any orthonormal
% basis~$\left\{ \ket{\psi_{j}}\right\} _{j=0}^{3}$, we can always
% pick 
% P_{1}=\proj{\psi_{0}}+\proj{\psi_{1}},\quad P_{2}=\proj{\psi_{0}}+\proj{\psi_{2}},\quad P_{3}=\proj{\psi_{0}}\textrm{ .}\end{array}
% With the probability measure~$\mu\left(P\right)=\melement{\psi}{P}$,
% where $\ket{\psi}=\frac{1}{2}\sum_{j=0}^{3}\ket{\psi_{j}}$, their
% probabilities are
% \mu\left(P_{1}\right)=1/2,\quad\mu\left(P_{2}\right)=1/2,\quad\mu\left(P_{3}\right)=1/4\textrm{ ,}\end{array}
% and we have
% \mu\left(P_{1}\right)\mu\left(P_{2}\right) & = & \frac{1}{4}=\mu\left(P_{3}\right)=\mu\left(P_{1}P_{2}\right)\textrm{ .}
% In particular, the following four Bell states form an orthonormal
% basis~\cite{CabelloPRL.86.2001}: 
% \ket{\phi^{\pm}} = \frac{1}{\sqrt{2}}\left(\ket{00}\pm\ket{11}\right),\quad
% \ket{\psi^{\pm}} = \frac{1}{\sqrt{2}}\left(\ket{01}\pm\ket{10}\right)\textrm{ .}
% Their sum is also a Bell state~
% \ket{\psi}=\frac{1}{2}\left(\ket{\phi^{+}}+\ket{\phi^{-}}+\ket{\psi^{+}}+\ket{\psi^{-}}\right)=\frac{1}{\sqrt{2}}\left(\ket{00}+\ket{01}\right)\textrm{ .}
% Thus, $\proj{\phi^{+}}+\proj{\psi^{+}}$ and $\proj{\phi^{+}}+\proj{\psi^{-}}$
% are independent projections with respect to the probability
% measure~$\mu\left(P\right)=\melement{\psi}{P}$.\qed\end{example}}

\footnote{\yutsung{The definition of independence is interesting, and we
definitely need to discuss it when we want to discuss Bell's theorem and the Kochen-Specker
theorem. However, the definition diverges. So we just leave it so far, and we
will go back if we really need it. (Maybe when discussing repeating
experiments?)}}

@@

\subsection{Measuring Quantum Probabilities}
\amr{preparation fuzzy, device fuzzy, Meyer~\cite{PhysRevLett.83.3751}}

\subsection{Quantum Interval-valued Probabilities Measures}
\cite{Vourdas2014, Vourdas2015}
